---
author: "Maks Nikiforov and Mark Austin"
date: "Due 10/31/2021"
bibliography: references.bib 
csl: apa.csl
params: 
      channel: ""
---


```{r image setup, include=FALSE}

##For markdown automation need a different image folder 
##  for each of the 6 channels
##Also setting up currentChannel variable 
if (params$channel=="data_channel_is_bus") {
  knitr::opts_chunk$set(fig.path = "images/bus/")
  currentChannel<-"Business"
} else if (params$channel=="data_channel_is_entertainment") {
  knitr::opts_chunk$set(fig.path = "images/entertainment/")
  currentChannel<-"Entertainment"
} else if (params$channel=="data_channel_is_lifestyle") {
  knitr::opts_chunk$set(fig.path = "images/lifestyle/")
  currentChannel<-"Lifestyle"
} else if (params$channel=="data_channel_is_socmed") {
  knitr::opts_chunk$set(fig.path = "images/socmed/")
  currentChannel<-"Social Media"
} else if (params$channel=="data_channel_is_tech") {
  knitr::opts_chunk$set(fig.path = "images/tech/")
  currentChannel<-"Tech"
} else if (params$channel=="data_channel_is_world") {
  knitr::opts_chunk$set(fig.path = "images/world/")
  currentChannel<-"World"
} 
#knitr::opts_chunk$set(fig.path = "images/")

```

---
title: "Analysis for the `r currentChannel` Channel"
---



```{r load packages,message=FALSE,echo=FALSE}
   library(tidyverse)
   library(caret)
   library(knitr)
   library(lares)
```

```{r import data, message=FALSE}
# Read all data into a tibble
fullData<-read_csv("./data/OnlineNewsPopularity.csv")

# Eliminate non-predictive variables
reduceVarsData<-fullData %>% select(-url,-timedelta)
#Are there other vars we do not need to use??

#test code to be removed later
#params$channel<-"data_channel_is_bus"

#filter by the current params channel
channelData<-reduceVarsData %>% filter(eval(as.name(params$channel))==1) 

###Can now drop the data channel variables 
channelData<-channelData %>% select(-starts_with("data_channel"))

```

## Introduction  

This page offers an exploratory data analysis of `r currentChannel` articles in the [online news popularity data set](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity). Two variables - `url` and `timedelta` - are non-predictive and have been removed. The remaining `r ncol(channelData)` variables comprise `r nrow(channelData)` observations, which makes up `r round(nrow(channelData)/nrow(fullData)*100, 1)` of the original data set.   

The broader purpose of this analysis is predicated on using supervised learning to predict a target variable - `shares`. To this end, the final sections outline four unique models for predicting the number of article shares and an assessment of their relative performance. Two models are rooted in multiple linear regression analysis, which assesses relationships between a response variable and two or more predictors. The remaining models are based on random forest and boosted tree techniques. The random forest method averages results from multiple decision trees which are fitted with a random parameter subset. The boosted tree method spurns averages in favor of results that stem from weighted iterations [@2021].             

## Summarizations  


### Numerical Summaries  

Summary information for shares grouped by whether an article was a weekend article or not. This gives an idea of the center and spread for shares.  

```{r shares summary}
channelData %>% group_by(is_weekend) %>% 
  summarise(Avg = mean(shares), Sd = sd(shares), 
    Median = median(shares), IQR =IQR(shares)) %>% kable(caption = "Summary Statistics for Shares")
```

The table below highlights variables with the highest and most significant correlations in the data set. This output may be considered when analyzing covariance to control for potentially confounding variables.  

```{r corr_cross, echo=TRUE, eval=TRUE, message=FALSE}
# Display top 10 highest correlations
covarianceDF <- corr_cross(df = channelData, max_pvalue = 0.05, top = 10, plot = 0) %>% 
  select(key, mix, corr, pvalue) %>% rename("Variable 1" = key, "Variable 2" = mix, 
                                            "Correlation" = corr, "p-value" = pvalue) 

# Display non-zero p-values
covarianceDF[4] <- format.pval(covarianceDF[4])

kable(covarianceDF)
```



### Contingency Tables  


```{r shares by size}
##I want to try and split up shares then create a contingency table


```


### Plots  

```{r graphOneA ,eval=FALSE}

###this one is in progress, hopefully a scatter plot
g<-ggplot(data = channelData,
          aes(x= avg_negative_polarity,y=shares))
g + geom_point(aes(color=as.factor(is_weekend))) +
  scale_y_continuous(trans = "pseudo_log")
  
  scale_y_continuous(labels = scales::comma) 



```


```{r histogram of shares}

###creating histogram of shares data 
g <- ggplot(channelData, aes( x = shares))
g + geom_histogram(binwidth=12000,color = "brown", fill = "green", 
  size = 1)  + labs(x="Article Shares", y="Pseudo Log of Count",
  title = "Histogram of Article Shares") +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0:3, 2000, 6000),minor_breaks = NULL) +
  scale_x_continuous(labels = scales::comma) 

```


```{r corrplot or ggpairs}

#depending on whether dr. post gives ok to do corrplot I'll do one of these for my 3rd graph

```



```{r barplot, echo=TRUE, eval=TRUE}
## Bar plot placeholder

# Subset columns to include only weekday_is_*
weekdayData <- channelData %>% select(starts_with("weekday_is"))

# Calculate sum of articles published in each week day
articlesPublished <- lapply(weekdayData, function(c) sum(c=="1"))

# Use factor to set specific order in bar plot
df <- data.frame(weekday=c("Monday", "Tuesday", "Wednesday", 
                           "Thursday", "Friday", "Saturday", "Sunday"),
                count=articlesPublished)
df$weekday = factor(df$weekday, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                           "Thursday", "Friday", "Saturday"))

# Create bar plot with total publications by day
weekdayBar <- ggplot(df, aes(x = weekday, y = articlesPublished)) + geom_bar(stat = "identity", color = "#123456", fill = "#0072B2") 
weekdayBar + labs(x = "Day", y = "Number published",
       title = "Article publications by day of week")
```




## Modeling  

```{r splitting data}

#Using set.seed per suggestion so that work will be reproducible
set.seed(20)

dataIndex <-createDataPartition(channelData$shares, p = 0.7, list = FALSE)

channelTrain <-channelData[dataIndex,]
channelTest <-channelData[-dataIndex,]

```


```{r train linear regression model one}
# Seeing "Error in summary.connection(connection) : invalid connection" after
# previous parallel computing runs, stopCluster(cl) may not be working as expected
cl <- makePSOCKcluster(6)
registerDoParallel(cl)

# Linear regression 
lmFit1 <- train(shares ~ ., data = channelTrain,
               method = "lm",
               preProcess = c("center", "scale"),
               trControl = trainControl(method = "cv", 
                                        number = 5))


stopCluster(cl)

# Predict using test data
predictLM1 <- predict(lmFit1, newdata = channelTest)

# Metrics
postResample(predictLM1, obs = channelTest$shares)

# Only RMSE
RMSE(channelTest$shares, predictLM1)
```


```{r train linear regression model two}

```


```{r train random forest model, eval=FALSE}

##without parallel code this was still running after 30 minutes so tried parallel next

##my pc has 8 cores so chose 5

##Followed Parallel instructions on caret page
##   https://topepo.github.io/caret/parallel-processing.html
##Even then it took 10 minutest to run
## and picked m=1 so not sure this is working correctly yet?

library(doParallel)
cl <- makePSOCKcluster(5)
registerDoParallel(cl)

rfFit <- train(shares ~ ., data = channelData,
               method = "rf",
               trControl = trainControl(method = "cv",
                                number = 5),
               tuneGrid = data.frame(mtry = 1:15))

stopCluster(cl)

rfFit

```


```{r train boosted tree model, include=FALSE}
# Seeing "Error in summary.connection(connection) : invalid connection"
# if I don't re-allocate cores for parallel computing
cl <- makePSOCKcluster(6)
registerDoParallel(cl)


# Boosted tree fit with tuneLength (let function decide parameter combinations)
boostedTreeFit <- train(shares ~ ., data = channelTrain,
               method = "gbm",
               preProcess = c("center", "scale"),
               trControl = trainControl(method = "cv", 
                                        number = 5),  
               tuneLength = 5)

# Define tuning parameters based on $bestTune from the permutations above
nTrees <- boostedTreeFit$bestTune$n.trees
interactionDepth = boostedTreeFit$bestTune$interaction.depth
minObs = boostedTreeFit$bestTune$n.minobsinnode
shrinkParam <- boostedTreeFit$bestTune$shrinkage

# Boosted tree fit with defined parameters
bestBoostedTree <- train(shares ~ ., data = channelTrain,
               method = "gbm",
               preProcess = c("center", "scale"),
               trControl = trainControl(method = "cv", 
                                        number = 5),  
               tuneGrid = expand.grid(n.trees = nTrees, interaction.depth = interactionDepth,
                                      shrinkage = shrinkParam, n.minobsinnode = minObs))


stopCluster(cl)

# summary(bestBoostedTree)

# Predict using test data
predictGBM <- predict(bestBoostedTree, newdata = channelTest)

# Metrics
postResample(predictGBM, obs = channelTest$shares)

# Only RMSE
RMSE(channelTest$shares, predictGBM)
```


## Model Comparisons  

This part needs to be automated.  Maybe create a function and iterate over these if they are similar?

```{r test linear regression model one}

```


```{r test linear regression model two}

```


```{r test random forest model}

```


```{r test boosted tree model}

```





