---
author: "Maks Nikiforov and Mark Austin"
date: "Due 10/31/2021"
bibliography: references.bib 
csl: apa.csl
params: 
      channel: ""
---


```{r image setup, include=FALSE}

##For markdown automation need a different image folder 
##  for each of the 6 channels
##Also setting up currentChannel variable 
if (params$channel=="data_channel_is_bus") {
  knitr::opts_chunk$set(fig.path = "images/bus/")
  currentChannel<-"Business"
} else if (params$channel=="data_channel_is_entertainment") {
  knitr::opts_chunk$set(fig.path = "images/entertainment/")
  currentChannel<-"Entertainment"
} else if (params$channel=="data_channel_is_lifestyle") {
  knitr::opts_chunk$set(fig.path = "images/lifestyle/")
  currentChannel<-"Lifestyle"
} else if (params$channel=="data_channel_is_socmed") {
  knitr::opts_chunk$set(fig.path = "images/socmed/")
  currentChannel<-"Social Media"
} else if (params$channel=="data_channel_is_tech") {
  knitr::opts_chunk$set(fig.path = "images/tech/")
  currentChannel<-"Tech"
} else if (params$channel=="data_channel_is_world") {
  knitr::opts_chunk$set(fig.path = "images/world/")
  currentChannel<-"World"
} 
#knitr::opts_chunk$set(fig.path = "images/")

```

---
title: "Analysis for the `r currentChannel` Channel"
---



```{r load packages,message=FALSE,echo=FALSE}
   library(tidyverse)
   library(caret)
   library(knitr)
   library(doParallel)
   library(lares)
   library(treemapify)
   library(rvest)
```

```{r import data, message=FALSE}
# Read all data into a tibble
fullData<-read_csv("./data/OnlineNewsPopularity.csv")

# Eliminate non-predictive variables
reduceVarsData<-fullData %>% select(-url,-timedelta)
#Are there other vars we do not need to use??

#test code to be removed later
#params$channel<-"data_channel_is_bus"

#filter by the current params channel
channelData<-reduceVarsData %>% filter(eval(as.name(params$channel))==1) 

# Temp, will clean up
channelDataURL <- fullData %>% filter(eval(as.name(params$channel))==1)

###Can now drop the data channel variables 
channelData<-channelData %>% select(-starts_with("data_channel"))

# Temp, will clean up
channelDataURL <- fullData %>% filter(eval(as.name(params$channel))==1)
```

## Introduction  

This page offers an exploratory data analysis of `r currentChannel` articles in the [online news popularity data set](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity). The top ten articles in this category, based on the number of shares on social media, include the following titles:  

```{r scrape titles, echo = FALSE, message=FALSE}
# Obtain top 10 rows by number of shares
topURL <- channelDataURL %>% slice_max(n = 10, shares) %>% select(url, shares) %>% 
  mutate(title = NA)

# Initialize vector to store formatted URLs
urlString <- vector()

# Populate "title" column with article titles
for (i in 1:nrow(topURL)) {
  urlString[i] <- paste0("", topURL$url[i], "")
  topURL$title[i] <- read_html(urlString[i]) %>% html_nodes("title") %>%
    html_text()
}

topURL %>% select(shares, title) %>% kable(col.names = c("Shares", "Article title"))
```


Two variables - `url` and `timedelta` - are non-predictive and have been removed. The remaining `r ncol(channelData)` variables comprise `r nrow(channelData)` observations, which makes up `r round(nrow(channelData)/nrow(fullData)*100, 1)` of the original data set. Fernandes et al., who sourced the data, concentrated on article characteristics such as verbosity and the polarity of content, publication day, the quantity of included media, and keyword attributes [@10.1007/978-3-319-23485-4_53]. A subset of these variables and the correlations between them are explored in subsequent sections.       

The broader purpose of this analysis is predicated on using supervised learning to predict a target variable - `shares`. To this end, the final sections outline four unique models for conducting such predictions and an assessment of their relative performance. Two models are rooted in multiple linear regression analysis, which assesses relationships between a response variable and two or more predictors. The remaining models are based on random forest and boosted tree techniques. The random forest method averages results from multiple decision trees which are fitted with a random parameter subset. The boosted tree method spurns averages in favor of results that stem from weighted iterations [@2021].             

## Summarizations  


### Numerical Summaries  

The first table summarizes information for article shares grouped by whether an article was a weekend article or not. This summary gives an idea of the center and spread of `shares` across type of day group levels.  

```{r shares weekend summary}
channelData %>% 
  mutate(dayType=ifelse(is_weekend,"Weekend","Weekday")) %>%
  group_by(dayType) %>% 
  summarise(Avg = mean(shares), Sd = sd(shares), 
    Median = median(shares), IQR =IQR(shares)) %>% kable()
```

The next tables gives expands on the idea of the first table by grouping `shares` by each day of the week.  This summary gives an idea of the center and spread of `shares` across day of the week group levels.  


```{r shares dow summary}
dowData<-channelData %>% select(starts_with("weekday_is"),shares) %>%
  mutate(dayofWeek=case_when(as.logical(weekday_is_monday)~"Monday",
                             as.logical(weekday_is_tuesday)~"Tuesday",
                             as.logical(weekday_is_wednesday)~"Wednesday",
                             as.logical(weekday_is_thursday)~"Thursday",
                             as.logical(weekday_is_friday)~"Friday",
                             as.logical(weekday_is_saturday)~"Saturday",
                             as.logical(weekday_is_sunday)~"Sunday")) %>%
  select(dayofWeek,shares)

dowLevels<-c("Monday","Tuesday","Wednesday",
             "Thursday","Friday","Saturday","Sunday")
dowData$dayofWeek<-factor(dowData$dayofWeek,levels = dowLevels)

dowData %>%  
  group_by(dayofWeek) %>% 
  summarise(Avg = mean(shares), Sd = sd(shares), 
    Median = median(shares), IQR =IQR(shares)) %>% kable()
```


The table below highlights variables with the highest and most significant correlations in the data set. This output may be considered when analyzing covariance to control for potentially confounding variables.  

```{r corr_cross, echo=TRUE, eval=TRUE, message=FALSE}
# Display top 10 highest correlations
covarianceDF <- corr_cross(df = channelData, max_pvalue = 0.05, top = 10, plot = 0) %>% 
  select(key, mix, corr, pvalue) %>% rename("Variable 1" = key, "Variable 2" = mix, 
                                            "Correlation" = corr, "p-value" = pvalue) 

# Display non-zero p-values
covarianceDF[4] <- format.pval(covarianceDF[4])

kable(covarianceDF)
```



### Contingency Tables  

The following contingency table displays counts and sums for the number of article shares within given ranges by the day of week shared.  Share ranges were selected to illustrate lower, medium, and higher ranges of shares.  Examining these counts can show possible patterns of shares by day or week and the range grouping for shares.  

```{r shares by size}

##dig.lab is needed to avoid R defaulting to scientific notation
kable(addmargins(table
                 (dowData$dayofWeek,cut(dowData$shares,
                  c(0,200,1000,10000,860000),dig.lab = 7))))



```


### Plots  

```{r graphOneA ,eval=FALSE}

###this one is in progress, hopefully a scatter plot
g<-ggplot(data = channelData,
          aes(x= avg_negative_polarity,y=shares))
g + geom_point(aes(color=as.factor(is_weekend))) +
  scale_y_continuous(trans = "pseudo_log")
  
  scale_y_continuous(labels = scales::comma) 



```

The following histogram looks at the distribution of `shares`.  A pseudo log y scale with modified y break values was used so that article `shares` with low frequency will appear.  We can tell from the histogram whether `shares` has a symmetric or skewed distribution.  The distribution is symmetric if the tails are the same around the center.  The distribution is right skewed if there is a long left tail and right skewed if there is a long right tail.  

```{r histogram of shares}

###creating histogram of shares data 
##scales comma was used to avoid the default scientific notation
##pseudo log with breaks was used to make low frequency values 
## more visisble
g <- ggplot(channelData, aes( x = shares))
g + geom_histogram(binwidth=12000,color = "brown", fill = "green", 
  size = 1)  + labs(x="Article Shares", y="Pseudo Log of Count",
  title = "Histogram of Article Shares") +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0:3, 2000, 6000),minor_breaks = NULL) +
  scale_x_continuous(labels = scales::comma) 

```


```{r corrplot or ggpairs}

#depending on whether dr. post gives ok to do corrplot I'll do one of these for my 3rd graph

```

```{r bar plot}
## Bar plot placeholder

# Subset columns to include only weekday_is_*
weekdayData <- channelData %>% select(starts_with("weekday_is"))

# Calculate sum of articles published in each week day
articlesPublished <- lapply(weekdayData, function(c) sum(c=="1"))

# Use factor to set specific order in bar plot
weekPubDF <- data.frame(weekday=c("Monday", "Tuesday", "Wednesday", 
                           "Thursday", "Friday", "Saturday", "Sunday"),
                count=articlesPublished)
weekPubDF$weekday = factor(weekPubDF$weekday, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                           "Thursday", "Friday", "Saturday"))

# Create bar plot with total publications by day
weekdayBar <- ggplot(weekPubDF, aes(x = weekday, y = articlesPublished)) + geom_bar(stat = "identity", color = "#123456", fill = "#0072B2") 
weekdayBar + labs(x = "Day", y = "Number published",
       title = "Article publications by day of week")
```

The box plot below examines the day of article publication (Monday-Sunday) and the associated distribution of article `shares`. The combination of a relatively higher median on Saturday and Sunday, as well as a comparably tall upper whisker, suggests that articles published on weekends may circulate more frequently on social media than articles published on weekdays.       

```{r boxpot preparation, echo=TRUE, eval=TRUE, message=FALSE, results='hide'}
# Subset columns to include only weekday_is_*, shares,
# create categorical variable, "day", denoting day of week (Mon-Sun)
medianShares <- channelData %>% select(starts_with("weekday_is"), shares) %>% mutate(day = NA)

# Populate "day"
for (i in 1:nrow(medianShares)) {
  if (medianShares$weekday_is_monday[i] == 1) {
    medianShares$day[i] = "Monday"
  }
  else if (medianShares$weekday_is_tuesday[i] == 1) {
    medianShares$day[i] = "Tuesday"
  }
  else if (medianShares$weekday_is_wednesday[i] == 1) {
    medianShares$day[i] = "Wednesday"
  }
  else if (medianShares$weekday_is_thursday[i] == 1) {
    medianShares$day[i] = "Thursday"
  }
  else if (medianShares$weekday_is_friday[i] == 1) {
    medianShares$day[i] = "Friday"
  }
  else if (medianShares$weekday_is_saturday[i] == 1) {
    medianShares$day[i] = "Saturday"
  }
  else if (medianShares$weekday_is_sunday[i] == 1) {
    medianShares$day[i] = "Sunday"
  }
  else {
    medianShares$day[i] = NA
  }
}

# Transform "day" into factor with levels to control order of boxplots
medianShares$day <- factor(medianShares$day, 
                           levels = c("Monday", "Tuesday", "Wednesday", 
                                      "Thursday", "Friday", "Saturday", "Sunday"))
```
```{r boxpot , echo=TRUE, eval=TRUE}
# Plot distribution of shares for each day of the week
sharesBox <- ggplot(medianShares, aes(x = day, y = shares, fill = day))

sharesBox + geom_boxplot(outlier.shape = NA) + 
  # Exclude extreme outliers, limit range of y-axis
  coord_cartesian(ylim = quantile(medianShares$shares, c(0.1, 0.95))) +
  # Remove legend after coloration
  theme(legend.position = "none") +
  labs(x = "Day", y = "Shares",
       title = "Distribution of article shares for each publication day") + scale_fill_brewer(palette = "Spectral")
```

For the empirical cumulative distribution function (ECDF) below, the `dplyr` ranking function `ntile()` divides `shares` into four groups. Observations with the fewest shares are placed into group 1, those with the most shares are placed into group 4, and intermediaries reside in groups 2 and 3. The horizontal axis lists word count, and the vertical axis lists the percentage of content with that word count. A divergence of the colored lines suggests that the number of words differs in content with the fewest and most shares.     

```{r ecdf, echo=TRUE, eval=TRUE, message=FALSE}
# Create variable to for binning the shares
binnedShares <- channelData %>% mutate(shareQuantile = ntile(channelData$shares, 4))
binnedShares <- binnedShares %>% mutate(totalMedia = num_imgs + num_videos)

# Render and label word count ECDF, group by binned shares
avgWordHisto <- ggplot(binnedShares, aes(x = n_tokens_content, colour = shareQuantile))
avgWordHisto + stat_ecdf(geom = "step", aes(color = as.character(shareQuantile))) +
  labs(title="ECDF - Number of words in the article \ngrouped by article shares (ranked)",
     y = "ECDF", x="Word count") + xlim(0,2000) + 
  scale_colour_brewer(palette = "Spectral", name = "Article shares \n(group rank)")
```


## Modeling  

```{r splitting data}

#Using set.seed per suggestion so that work will be reproducible
set.seed(20)

dataIndex <-createDataPartition(channelData$shares, p = 0.7, list = FALSE)

channelTrain <-channelData[dataIndex,]
channelTest <-channelData[-dataIndex,]

```


```{r train linear regression model one}
# Seeing "Error in summary.connection(connection) : invalid connection" after
# previous parallel computing runs, stopCluster(cl) may not be working as expected
cl <- makePSOCKcluster(6)
registerDoParallel(cl)

# Linear regression 
lmFit1 <- train(shares ~ ., data = channelTrain,
               method = "lm",
               preProcess = c("center", "scale"),
               trControl = trainControl(method = "cv", 
                                        number = 5))


stopCluster(cl)

# Predict using test data
predictLM1 <- predict(lmFit1, newdata = channelTest)

# Metrics
postResample(predictLM1, obs = channelTest$shares)

# Only RMSE
# RMSE(channelTest$shares, predictLM1)
```


```{r train linear regression model two}

```


```{r train random forest model, eval=FALSE}

##without parallel code this was still running after 30 minutes so tried parallel next

##my pc has 8 cores so chose 5

##Followed Parallel instructions on caret page
##   https://topepo.github.io/caret/parallel-processing.html
##Even then it took 10 minutest to run
## and picked m=1 so not sure this is working correctly yet?

cl <- makePSOCKcluster(5)
registerDoParallel(cl)

rfFit <- train(shares ~ ., data = channelData,
               method = "rf",
               trControl = trainControl(method = "cv",
                                number = 5),
               tuneGrid = data.frame(mtry = 1:15))

stopCluster(cl)

rfFit

```


```{r train boosted tree model}
# Seeing "Error in summary.connection(connection) : invalid connection"
# if I don't re-allocate cores for parallel computing
cl <- makePSOCKcluster(6)
registerDoParallel(cl)


# Boosted tree fit with tuneLength (let function decide parameter combinations)
boostedTreeFit <- train(shares ~ ., data = channelTrain,
               method = "gbm",
               preProcess = c("center", "scale"),
               trControl = trainControl(method = "cv", 
                                        number = 5),  
               tuneLength = 5)

# Define tuning parameters based on $bestTune from the permutations above
nTrees <- boostedTreeFit$bestTune$n.trees
interactionDepth = boostedTreeFit$bestTune$interaction.depth
minObs = boostedTreeFit$bestTune$n.minobsinnode
shrinkParam <- boostedTreeFit$bestTune$shrinkage

# Boosted tree fit with defined parameters
bestBoostedTree <- train(shares ~ ., data = channelTrain,
               method = "gbm",
               preProcess = c("center", "scale"),
               trControl = trainControl(method = "cv", 
                                        number = 5),  
               tuneGrid = expand.grid(n.trees = nTrees, interaction.depth = interactionDepth,
                                      shrinkage = shrinkParam, n.minobsinnode = minObs))


stopCluster(cl)

# summary(bestBoostedTree)

# Predict using test data
predictGBM <- predict(bestBoostedTree, newdata = channelTest)

# Metrics
postResample(predictGBM, obs = channelTest$shares)

# Only RMSE
# RMSE(channelTest$shares, predictGBM)
```


## Model Comparisons  

This part needs to be automated.  Maybe create a function and iterate over these if they are similar?

```{r test linear regression model one}

```


```{r test linear regression model two}

```


```{r test random forest model}

```


```{r test boosted tree model}

```

## References 



